---
title: "Spotify"
author: "Daniel Zapata"
date: "6 de marzo de 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# loading required packages
```{r}
library(data.table)
# library(tm)
library(stringi)
library(spotifyr)
library(plyr)
library(tidyverse)
library(httr)
library(rvest)
library(stringr)
library(ggthemes)
library(tidytext)
library(wordcloud)
library(ggridges)
library(wesanderson)
library(yarrr)
library(knitr)
library(kableExtra)
library(radarchart)

```

# set up Spotify client ID and client secret
```{r}
creds <- readLines("spotify_credentials.txt")
Sys.setenv(SPOTIFY_CLIENT_ID = creds[1])
Sys.setenv(SPOTIFY_CLIENT_SECRET = creds[2])
```

#using spoitfyr
```{r}
fellas <- get_artist_audio_features("losPetitFellas") %>% data.table()
fellas
```

```{r}
unique(fellas$album_name)
fellas[, .N, by = album_name]
```

En Genius no se encuentra el album 1320 Sijazz que es un album en vivo. Tampoco estan todas las canciones de souvenir ya que algunas se repiten de albunes anteriores. 

# Getting artist ID on Genius
```{r}
creds <- readLines("genius_credentials.txt")
token <- creds[3]

genius_get_artists <- function(artist_name, n_results = 10) {
  baseURL <- 'https://api.genius.com/search?q=' 
  requestURL <- paste0(baseURL, gsub(' ', '%20', artist_name),
                       '&per_page=', n_results,
                       '&access_token=', token)
  
  res <- GET(requestURL) %>% content %>% .$response %>% .$hits
  
  map_df(1:length(res), function(x) {
    tmp <- res[[x]]$result$primary_artist
    list(
      artist_id = tmp$id,
      artist_name = tmp$name
    )
  }) %>% unique
}


genius_artists <- genius_get_artists('LosPetitFellas')
genius_artists
```

# Getting track urls
```{r}
baseURL <- 'https://api.genius.com/artists/' 
requestURL <- paste0(baseURL, genius_artists$artist_id[1], '/songs')

track_lyric_urls <- list()
i <- 1
while (i > 0) {
  tmp <- GET(requestURL, query = list(access_token = token, per_page = 50, page = i)) %>% content %>% .$response
  track_lyric_urls <- c(track_lyric_urls, tmp$songs)
  if (!is.null(tmp$next_page)) {
    i <- tmp$next_page
  } else {
    break
  }
}



# Filtering to get urls only for tracks on which fellas Swift is the primary artist
filtered_track_lyric_urls <- c()
filtered_track_lyric_titles <- c()
index <- c()

for (i in 1:length(track_lyric_urls)) {
  if (track_lyric_urls[[i]]$primary_artist$name == "LosPetitFellas") {
    filtered_track_lyric_urls <- append(filtered_track_lyric_urls, track_lyric_urls[[i]]$url)
    filtered_track_lyric_titles <- append(filtered_track_lyric_titles, track_lyric_urls[[i]]$title)
    
    index <- append(index, i)
  }
}


```

hay que juntar ambas funtes, para que sea más facil eliminaré signos de puntuación y tildes de los titulos
```{r}
filtered_track_lyric_titles <- tolower(filtered_track_lyric_titles)
filtered_track_lyric_titles <- 
  stri_trans_general(filtered_track_lyric_titles,"Latin-ASCII")
filtered_track_lyric_titles <- tm::removePunctuation(filtered_track_lyric_titles)
filtered_track_lyric_titles
```

```{r}
fellas$track_name <-  tolower(fellas$track_name)
fellas$track_name <- stri_trans_general(fellas$track_name,"Latin-ASCII")
fellas$track_name <- tm::removePunctuation(fellas$track_name)
fellas$track_name
```

```{r}
fellas_lyrics <- data.frame(filtered_track_lyric_urls, filtered_track_lyric_titles)
fellas_lyrics <- fellas_lyrics[filtered_track_lyric_titles %in% fellas$track_name, ]
```

```{r}
filtered_track_lyric_titles[!(filtered_track_lyric_titles %in% fellas$track_name)]
```

```{r}
fellas$track_name[!(fellas$track_name %in% filtered_track_lyric_titles)]
```
```{r}
fellas_lyrics$filtered_track_lyric_urls <- as.character(fellas_lyrics$filtered_track_lyric_urls)
fellas_lyrics$filtered_track_lyric_titles <- as.character(fellas_lyrics$filtered_track_lyric_titles)
```

# Webscraping lyrics using rvest 
```{r}
lyric_text <- c()
for (i in 1:nrow(fellas_lyrics)) {
  lyric_text_pivot <- read_html(fellas_lyrics$filtered_track_lyric_urls[i]) %>% 
    html_nodes(".lyrics p") %>% 
    html_text()
  lyric_text[i] <- lyric_text_pivot[lyric_text_pivot != ""]
}

```

# Cleaning and standardizing lyrics
```{r}
for (i in 1:nrow(fellas_lyrics)) {
  lyric_text[i] <- gsub("([a-z])([A-Z])", "\\1 \\2", lyric_text[i])
  lyric_text[i] <- gsub("\n", " ", lyric_text[i])
  lyric_text[i] <- gsub("\\[.*?\\]", " ", lyric_text[i])
  lyric_text[i] <- tolower(lyric_text[i])
  lyric_text[i] <- gsub("[ [:punct:] ]", " ", lyric_text[i])
  lyric_text[i] <- gsub(" {2,}", " ", lyric_text[i])
  lyric_text[i] <-  stri_trans_general(lyric_text[i],"Latin-ASCII")
   lyric_text[i] <- tm::removePunctuation(lyric_text[i])
}
lyric_text[1]
```

```{r}
genius_data <- data.frame(track_name = fellas_lyrics$filtered_track_lyric_titles, lyrics = lyric_text)
genius_data$track_name <- as.character(genius_data$track_name)
genius_data$lyrics <- as.character(genius_data$lyrics)
genius_data
```

# joining Spotify and Genius data
```{r}
spotify_genius <- full_join(genius_data, fellas, by = "track_name")
spotify_genius
```

```{r}
spotify_genius <- data.table(spotify_genius)
setkey(spotify_genius, "album_release_year")
spotify_genius[, album_name := factor(album_name, ordered = T)]
spotify_genius[, valence := as.numeric(valence)]
options("digits" = 3)

```

```{r}

# valence ridge plot (I used fig.height = 6, fig.width = 6 in an rmd)
spotify_genius %>% ggplot(aes(x = valence, fill = album_name)) + 
  geom_density(position = "identity", alpha = 0.6) +
   scale_x_continuous(name = "Valencia",
                           breaks = seq(0, 1, 0.05),
                           limits=c(0, 1)) +
        scale_y_continuous(name = "Density") +
        ggtitle("Distribución de la valencia por album") +
        theme_bw() +
        theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma")) +
        scale_fill_brewer(palette="Accent")

```
# table: album by mean valence
```{r}
spotify_genius %>% 
  group_by(album_name) %>% 
  summarise(mean(valence)) %>% 
  arrange(desc(`mean(valence)`)) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F, position = "left") %>% 
  row_spec(row = 1:4, background = "#fffce4", color = "red")
```

# table: top 5 songs by valence
```{r}
spotify_genius %>% 
  select(track_name, album_name, valence) %>% 
  top_n(5) %>% 
  arrange(-valence) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F, position = "left") %>% 
  row_spec(row = 1:5, background = "azure", color = "deeppink")

```
# table: top 5 songs by lower valence
```{r}
spotify_genius %>% 
  select(track_name, album_name, valence) %>% 
  top_n(-5) %>% 
  arrange(-valence) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F, position = "left") %>% 
  row_spec(row = 1:5, background = "azure", color = "deeppink")

```
# sonic score graph
```{r}
pirateplot(valence + danceability + energy ~ album_release_date, spotify_genius,
           pal = c(wes_palettes$GrandBudapest2, wes_palettes$Moonrise3[1:2]), 
           xlab = "álbum", ylab = "sonic score",
           theme = 0, point.o = 0.7, avg.line.o = 1, jitter.val = .05, 
           bty = "n", cex.axis = 0.6, xaxt = "n") 
axis(1, cex.axis = 0.6, lwd = 0)
legend("topright", c("1:Historias Mínimas", "2: Souvenir", "3: Formas Para Perderse o I.D.E.A.S", "4: 1320 Sijazz"), bty = "n", cex = 0.6) 
```
```{r}
spotify_genius %>% 
  mutate(sonic_score = valence + danceability + energy) %>% 
  select(album_name, track_name, sonic_score) %>% 
  arrange(desc(sonic_score)) %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left") %>% 
  row_spec(row = 1:46, background = "seashell", color = "#b39db2")
```

# album by danceability
```{r}
spotify_genius %>% 
  group_by(album_name) %>% 
  summarise(mean(danceability)) %>% 
  arrange(desc(`mean(danceability)`)) %>% 
  kable() %>% 
  kable_styling(full_width = F, position = "left") %>% 
  row_spec(row = 1:4, background = "seashell", color = "#b39db2")
```

# Lyrics 
```{r}
tidy_fellas <- spotify_genius %>% unnest_tokens(word, lyrics)

custom_stop_words <- bind_rows(stop_words,
                               data_frame(
                               word = stri_trans_general(tm::stopwords("spanish") , "Latin-ASCII"),
                               lexicon = "custom"
                               ))

tidier_fellas  <- tidy_fellas  %>%
    anti_join(custom_stop_words) 

word_count <- tidier_fellas %>%
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>%
  ungroup()


```

```{r}
wordcloud(words = word_count$word, freq = word_count$n,
          max.words=100, random.order=FALSE, 
          colors= c(wes_palettes$Moonrise3[c(1:2,5)], wes_palettes$Royal2[5]))
```

```{r}
# how many tracks does the word "remember" appear in?
tidier_fellas %>% 
  select(track_name, word) %>% 
  filter(word == "solo") %>% 
  unique() %>% 
  select(track_name)
```
```{r}
# wordcloud: fellas Swift the album
word_count_ts <- tidier_fellas %>%
  filter(album_name == "Historias Mínimas") %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>%
  ungroup()

wordcloud(words = word_count_ts$word, freq = word_count_ts$n,
          max.words=25, random.order=FALSE, 
          colors= c(wes_palettes$GrandBudapest2[3:1]))
```

```{r}
# wordcloud: reputation
word_count_rep <- tidier_fellas %>%
  filter(album_name == "Formas Para Perderse o I.D.E.A.S") %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>%
  ungroup()

wordcloud(words = word_count_rep$word, freq = word_count_rep$n,
          max.words=25, random.order=FALSE, 
          colors= c(wes_palettes$GrandBudapest2[3:1]))
```

# more cleaning, can be done earlier
```{r}
tidier_fellas$album_release_year <- as.character(tidier_fellas$album_release_year)
tidier_fellas$album_release_year <- as.numeric(substr(tidier_fellas$album_release_year, 1, 4))

tidy_fellas$album_release_year <- as.character(tidy_fellas$album_release_year)
tidy_fellas$album_release_year <- as.numeric(substr(tidy_fellas$album_release_year, 1, 4))
```

# creating a "lexical diversity" dataset
```{r}
unique(tidy_fellas$album_release_date)
```


```{r}
tidy_fellas <- data.table(tidy_fellas)
lexical_diversity <-
tidy_fellas[is.na(word) == F, .(lex_div = uniqueN(word) / length(word)), by = .(track_name, album_name)]
levels(lexical_diversity$album_name)
```

# lexical diversity plot
```{r}
pirateplot(lex_div ~ album_name, lexical_diversity,
           pal = c("cyan3", "darkgoldenrod1", "maroon4", "red3", "#b39db2", "black"),
           xlab = "album", ylab = "lexical diversity",
           theme = 0, point.o = 0.5, avg.line.o = 1, jitter.val = .05, 
           bty = "n", cex.axis = 0.6, xaxt = "n") 
axis(1, cex.axis = 0.6, lwd = 0)
legend("topright", c("1: Formas Para Perderse o I.D.E.A.S","2:Historias Mínimas", "3: Souvenir"), bty = "n", cex = 0.6) 

```

```{r}

```

```{r}

```

